---
title: "Linear Regression Model to predict Diamond Price"
author: "Dr Alina"
date: "September 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading The Diamond Data

```{r}
library(ggplot2)
data(diamonds)
str(diamonds)
```

###Scatter Plot Review
The scatterplot is a powerful tool to help you understand
the relationship between two continuous variables.

```{r Scatterplot Review}
# Let's start by examining two variables in the data set.
# We can quickly see if the relationship is linear or not.
# In this case, we can use a variety of diamond
# characteristics to help us figure out whether
# the price advertised for any given diamond is 
# reasonable or a rip-off.

# Let's consider the price of a diamond and it's carat weight.
# Create a scatterplot of price (y) vs carat weight (x).

# Limit the x-axis and y-axis to omit the top 1% of values.

# ENTER YOUR CODE BELOW THIS LINE
# ================================================================

ggplot(diamonds,(aes(x = carat, y = price)))+
  geom_point(fill = I('#F79420'),color = I('red'),shape= 21)+
  xlim(0,quantile(diamonds$carat,.99))+
  ylim(0,quantile(diamonds$price,.99))
```

***

### Price and Carat Relationship
Response:

1. there is non linear relationship between both variables, maybe its exponential or else.

2. Variance or dispersion increases as the carat size increases.

***
```{r To verify that linear line is not fit}
ggplot(diamonds,(aes(x = carat, y = price)))+
  geom_point(fill = I('#F79420'),color = I('red'),shape= 21)+
  stat_smooth(method = lm)+
  xlim(0,quantile(diamonds$carat,.99))+
  ylim(0,quantile(diamonds$price,.99))
```


### ggpairs Function
Notes:

https://s3.amazonaws.com/udacity-hosted-downloads/ud651/ggpairs_landscape.pdf

```{r ggpairs Function, echo = FALSE,include =FALSE}
# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
#set.seed(20022012)
#diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
#ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))
```

### The Demand of Diamonds
Notes: using log scale for price

```{r The Demand of Diamonds}
# Create two histograms of the price variable
# and place them side by side on one output image.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.

library(gridExtra)

p1 <- ggplot(diamonds,(aes(x = price)))+
  geom_histogram(binwidth = 100, fill = I('#099DD9'))+
  ggtitle('Diamond prices')

p2 = ggplot(diamonds,(aes(x = log10(price))))+
  geom_histogram(binwidth = 0.05, fill = I('#F79420'))+
  ggtitle('log Price')


grid.arrange(p1,p2,ncol = 2)
```


The two peaks shows the type of buyer as expected. 
 rich buyer and poor buyer.
  log 10 scale has bimodal and bell curve type of distribution, which is better in behavior.
***

### Connecting Demand and Price Distributions
Notes:
 On the log scale, the
prices look less dispersed at the high end of Carat
size and price, but actually we can do better. Let's
try using the cube root of Carat in light of our speculation about flaws being
exponentially more likely in diamonds with more volume
***

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(diamonds,(aes(x = carat, y = log10(price))))+
  geom_point(fill = I('#F79420'),color = I('brown'),shape= 21)+
  ggtitle('Carat vs (log10)Prices')
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', 
                                      transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```



#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point()+
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

*With these transformations
that we used to get our data on this nice scale. Things look
almost linear. We can now move forward and see about
modelling our data using just a linear model.*

***

### Overplotting Revisited

When multiple points are taken on same value.
to correct them

1. add transperency

2. add jitter points, adjust the shape

3. additional info by adding layer of geom_smooth or geom_quantile or geom_density_2D

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = TRUE))

head(sort(table(diamonds$price), decreasing = TRUE))
```


```{r Overplotting Revisited}
# Add a layer to adjust the features of the
# scatterplot. Set the transparency to one half,
# the size to three-fourths, and jitter the points.

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 0.75, position ='jitter') + 
  geom_count()+
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***
```{r, echo=FALSE}
library(RColorBrewer)
```

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
# Adjust the code below to color the points by clarity.

# A layer called scale_color_brewer() has 
# been added to adjust the legend and
# provide custom colors.

ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```



### Price vs. Carat and Color


Alter the code below.
```{r Price vs. Carat and Color}
# Adjust the code below to color the points by cut.
# Change any other parts of the code as needed.

ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:
yes . a pattern is observed in the color and price relation.

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:
sdigits = 3
This will ensure that the output at the end of the table shows three significant digits like shown in the video.

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5,sdigits = 3)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***
```{r}
load("BigDiamonds.Rda")
```

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}
diamondsbig$logprice <- log(diamondsbig$price)
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), 
         data = diamondsbig[diamondsbig$price < 10000
                            & diamondsbig$cert == "GIA",])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5,sdigits = 3)



# DO NOT ALTER THE CODE BELOW THIS LINE (Tables your models and pulls out the statistics)
#===========================================================================================
#suppressMessages(library(lattice))
#suppressMessages(library(MASS))
#suppressMessages(library(memisc))
models <- mtable(m1, m2, m3, m4, m5)
```


## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***
